---
title: "Reproducible science workshop"
subtitle: "Exercise 1 data manipulation Solutions"
author: "[**_Gbadamassi G.O. Dossa_**](https://scholar.google.com/citations?user=w20AkZcAAAAJ&hl=en)"
date: "last updated: `r Sys.Date()`"
output:
  xaringan::moon_reader:
    css: [default, "slides-theme.css"]
    lib_dir: libs
    nature:
      #ratio: '16:9'
      highlightStyle: github
      highlightLines: true
      slideNumberFormat: ''
      titleSlideClass: [center, middle]
---

```{r setup, include=FALSE, echo=FALSE}
#options(htmltools.dir.version = FALSE)
#knitr::opts_chunk$set(comment = "", warning = FALSE, message = FALSE)
knitr::opts_chunk$set(warning = FALSE, message = FALSE)
library(palmerpenguins)
library(tidyverse)
```

## Question 1a 
a. Display the data `penguins`.

.tiny-font[
```{r}
penguins # display data
```
]

---

## Question 1a 

.tiny-font[
```{r}
penguins %>% glimpse() # display data
```
]

---

## Question 1b
b. Make sure you understand the columns we have in this dataset.  

c. Filter out penguins for which sex is missing.  

d. Select variables species, island, bill_length_mm and body_mass_g.  

e. Store the new dataset in a dat object.  


```{r}
dat <- penguins %>%
# filter out missing sex
  filter(!is.na(sex)) %>% 
# select variables
  select(species, island, bill_length_mm, body_mass_g) 
```

---

## Question 2a
a. How many penguins do we have in the dataset?  



.tiny-font[
```{r}
dat # nb of penguins
```
]

---

## Questions 2b and 2d
b. How many species?   

d. Count the number of penguins per species.   


```{r}
# nb of species, and penguins per species
dat %>% count(species) 
```

---

## Questions 2b and 2d

```{r}
dat %>% count(species, sort = TRUE) # idem, arranged by n
```


---

## Question 2c
c. How many islands?   

```{r}
dat %>% count(island) # nb of island, and penguins per island
```


---

## Question 2e
e. Count the number of penguins per species and per island.    

```{r}
# penguins per species and island
dat %>% count(species, island) 
```



---

## Question 3a: mean body mass

.tiny-font[
```{r}
# option 1
dat %>%
  mutate(mean_bm = mean(body_mass_g))
```
]

---

## Question 3a: mean body mass
a. Calculate the overall mean body mass.  


```{r}
# option 2
dat %>%
  summarise(mean_bm = mean(body_mass_g))
```


---

## Question 3b: mean body mass per species
b. Calculate the mean body mass for each species.  


.tiny-font[
```{r}
# option 1
dat %>%
  group_by(species) %>% 
  mutate(mean_bm = mean(body_mass_g))
```
]



---

## Question 3b: mean body mass per species
b. Calculate the mean body mass for each species.  

```{r}
# option 2
dat %>% 
  group_by(species) %>% 
  summarize(mean_bm = mean(body_mass_g))
```


---

## Question 3c: mean traits
c. Calculate the mean of both traits bill length and body mass measured for each species.   

```{r}
# all at once, through column selection
dat %>% 
  group_by(species) %>% 
  summarize(across(bill_length_mm:body_mass_g, mean))
```



---

## Question 3c: mean traits
c. Calculate the mean of both traits bill length and body mass measured for each species.   

```{r}
# all at once, through column format selection
dat %>% 
  group_by(species) %>% 
  summarize(across(where(is.numeric), mean))
```
