---
title: "Exercice 1 Data manipulation solution"
author: "Gbadamassi G.O. Dossa"
date: "19/10/2021"
output:
#rmdformats::material
  html_document:
    fig_caption: yes
    theme: spacelab #readable #sandstone #spacelab #flatly
    highlight: pygments #tango #kate
    toc: TRUE
    toc_depth: 3
    number_sections: TRUE
    toc_float:
      collapsed: TRUE
      smooth_scroll: TRUE
      toc: true
---

```{r setup, include=FALSE, echo=FALSE}
#options(htmltools.dir.version = FALSE)
#knitr::opts_chunk$set(comment = "", warning = FALSE, message = FALSE)
knitr::opts_chunk$set(warning = FALSE, message = FALSE)
library(palmerpenguins)
library(tidyverse)
```

# Question 1a-a 

Display the overview of the data penguins
```{r}
penguins # display data
```



# Question 1a-b


```{r}
penguins %>% # This %>% is for then  
  glimpse() # display data

summary(penguins)
```

# Question 1b

```{r}
dat <- penguins %>%
# select variables
  select(species, island, bill_length_mm, body_mass_g) 
summary(dat)
```

If you want to deal with *NA* directly you can do as follows
```{r}
dat_billmean<-dat%>%
  group_by(species)%>%
  summarise(mean_bill=mean(bill_length_mm, na.rm=TRUE))
dat_billmean
```


```{r}

dat_na<-penguins %>%
  filter(!is.na(sex))
dim(dat_na)
dat <- penguins %>%
# filter out missing sex
  filter(!is.na(sex)) %>% 
# select variables
  select(species, island, bill_length_mm, body_mass_g) 
dat
```

# Question 2a


```{r}
dat # nb of penguins
```

# Questions 2b and 2d

```{r}
# nb of species, and penguins per species
dat %>% count(species) 
```


# Questions 2b and 2d

```{r}
dat %>% count(species, sort = TRUE) # idem, arranged by n
```

# Question 2c

```{r}
dat %>% count(island) # nb of island, and penguins per island
```

# Question 2e

```{r}
# penguins per species and island
dat %>% count(species, island) 
```


# Question 3a: mean body mass


```{r}
# option 1
dat %>%
  mutate(mean_bm = mean(body_mass_g))
```


# Question 3a: mean body mass

```{r}
# option 2
dat %>%
  summarise(mean_bm = mean(body_mass_g))
```

# Question 3b: mean body mass per species


```{r}
# option 1
dat %>%
  group_by(species) %>% 
  mutate(mean_bm = mean(body_mass_g))
```


# Question 3b: mean body mass per species

```{r}
# option 2
dat %>% 
  group_by(species) %>% 
  summarize(mean_bm = mean(body_mass_g))
```


# Question 3c: mean traits

```{r}
# all at once, through column selection
dat %>% 
  group_by(species) %>% 
  summarize(across(bill_length_mm:body_mass_g, mean))
```




# Question 3c: mean traits

```{r}
# all at once, through column format selection

dat %>% 
  group_by(species) %>% 
  summarize(across(where(is.numeric), mean))
```
