---
ttitle: "Reproducible science workshop"
subtitle: "Exercise 2 Data visualization: ggplot2-solutions"
date: (Created on Sept 31 2021 and last updated on `r format(Sys.time(), "%a %b %d %Y %H:%M:%S")`)
output: 
#rmdformats::material
  html_document:
    fig_caption: yes
    theme: spacelab #readable #sandstone #spacelab #flatly
    highlight: pygments #tango #kate
    toc: TRUE
    toc_depth: 3
    number_sections: TRUE
    toc_float:
      collapsed: TRUE
      smooth_scroll: TRUE
      toc: true
---

```{r setup, include=FALSE, echo=FALSE}
#options(htmltools.dir.version = FALSE)
#knitr::opts_chunk$set(comment = "", warning = FALSE, message = FALSE)
knitr::opts_chunk$set(warning = FALSE, message = FALSE)
library(palmerpenguins)
library(tidyverse)
```

# Motivation

We use data from package `palmerpenguins` to familiarize ourselves with data visualization with the tidyverse. We first load the packages we will need. 
```{r}
library(palmerpenguins)
library(tidyverse)
```


# Question 1a: histogram of body mass
a. Build a histogram of body mass.   

```{r}
penguins%>%
  glimpse()
```

```{r question1a, eval = FALSE}
penguins %>%
  ggplot() +
  aes(x = body_mass_g) + 
  geom_histogram()
```

```{r question1a, echo = FALSE, fig.width = 5.5, fig.asp = 0.618, dev = "svg"}
```


# Question 1b: a color per species
b. Fill in histogram with a color per species.      

```{r question1b, eval = FALSE}
penguins %>%
  ggplot() +
  aes(x = body_mass_g) + 
  geom_histogram(aes(fill = species))
```

```{r question1b, echo = FALSE, fig.width = 5.5, fig.asp = 0.618, dev = "svg"}
```


# Question 1c: darkorange, purple and cyan4
c. Change the color by default to darkorange, purple and cyan4.    

```{r question1c, eval = FALSE}
penguins %>%
  ggplot() +
  aes(x = body_mass_g) + 
  geom_histogram(aes(fill = species)) +
  scale_fill_manual(values = c("darkorange","purple","cyan4")) 
```


```{r question1c, echo = FALSE, fig.width = 5.5, fig.asp = 0.618, dev = "svg"}
```


# Question 1d: adjust transparency

d. Adjust transparency for all three histograms.     

```{r question1d, eval = FALSE}
penguins %>%
  ggplot() +
  aes(x = body_mass_g) + 
  geom_histogram(aes(fill = species), alpha = 0.5) +
  scale_fill_manual(values = c("darkorange","purple","cyan4")) 
```

```{r question1d, echo = FALSE, fig.width = 5.5, fig.asp = 0.618, dev = "svg"}
```


# Question 1e: add titles
e. Add a title to the axes, the legend and the figure.    


```{r question1e, eval = FALSE}
penguins %>%
  ggplot() +
  aes(x = body_mass_g) + 
  geom_histogram(aes(fill = species), 
                 alpha = 0.5) +
  scale_fill_manual(values = c("darkorange",
                               "purple",
                               "cyan4")) +
  labs(x = "Body mass (g)",
       y = "Frequency",
       title = "Penguin body mass",
       fill = "Penguin species")
```


```{r question1e, echo = FALSE, fig.width = 5.5, fig.asp = 0.618, dev = "svg"}
```



# Question 1f: change theme
f. Change the theme used by default.    

```{r question1f, eval = FALSE}
penguins %>%
  ggplot() +
  aes(x = body_mass_g) + 
  geom_histogram(aes(fill = species), 
                 alpha = 0.5) +
  scale_fill_manual(values = c("darkorange",
                               "purple",
                               "cyan4")) +
  labs(x = "Body mass (g)",
       y = "Frequency",
       title = "Penguin body mass",
       fill = "Penguin species") +
  theme_minimal()
```


```{r question1f, echo = FALSE, fig.width = 5.5, fig.asp = 0.618, dev = "svg"}
```

# Question 1g: split by sex
g. Do a-f for males and females and display on same figure.    

```{r question1g, eval = FALSE}
penguins %>%
# filter out penguins w/ missing sex
  filter(!is.na(sex)) %>% 
  ggplot() +
  aes(x = body_mass_g) + 
  geom_histogram(aes(fill = species),
# fill histogram, adjust transparency                 
                 alpha = 0.5) + 
# change colour
  scale_fill_manual(values = c("darkorange",
                               "purple",
                               "cyan4")) + 
  labs(x = "Body mass (g)", # x lab
       y = "Frequency", # y lab
       title = "Penguin body mass", 
       fill = "Penguin species") + 
  theme_minimal() + # change theme
# hist by sex, w/ diff X scale
  facet_wrap(~sex, scales = "free_x") 
```
There is NA in sex so we need to filter our them before plotting.

```{r , eval = FALSE}
penguins %>%
# filter out penguins w/ missing sex
  filter(!is.na(sex)) %>% 
  ggplot() +
  aes(x = body_mass_g) + 
  geom_histogram(aes(fill = species),
# fill histogram, adjust transparency                 
                 alpha = 0.5) + 
# change colour
  scale_fill_manual(values = c("darkorange",
                               "purple",
                               "cyan4")) + 
  labs(x = "Body mass (g)", # x lab
       y = "Frequency", # y lab
       title = "Penguin body mass", 
       fill = "Penguin species") + 
  theme_minimal() + # change theme
# hist by sex, w/ diff X scale
  facet_wrap(~sex, scales = "free_x") 
```


```{r, echo = FALSE, fig.width = 5.5, fig.asp = 0.618, dev = "svg"}
```


# Question 2a: scatter plot
a. Build a scatter plot of body mass with respect to flipper length.      


```{r question2a, eval = FALSE}
penguins %>% 
  ggplot() + 
  aes(x = flipper_length_mm,
      y = body_mass_g) +
  geom_point() +
  theme_bw()
```



```{r question2a, echo = FALSE, fig.width = 5.5, fig.asp = 0.618, dev = "svg"}
```


# Question 2b: species-specific shapes
b. Consider a shape for the points by allowing different species with d. Increase point size and adjust transparency. 


```{r question2b, eval = FALSE}
penguins %>% 
  ggplot() + 
  aes(x = flipper_length_mm,
      y = body_mass_g) +
  geom_point(aes(shape = species),
             size = 3,
             alpha = 0.5) +
  theme_minimal()
```


```{r question2b, echo = FALSE, fig.width = 5.5, fig.asp = 0.618, dev = "svg"}
```

# Question 2c: species-specific colors
c. Consider a color for the points different for each species. Increase point size and adjust transparency.   

```{r question2c, eval = FALSE}
penguins %>% 
  ggplot() + 
  aes(x = flipper_length_mm,
      y = body_mass_g) +
  geom_point(aes(color = species),
             size = 3,
             alpha = 0.8) +
  theme_minimal()
```

Different shapes and different colours
```{r , eval = FALSE}
penguins %>% 
  ggplot() + 
  aes(x = flipper_length_mm,
      y = body_mass_g) +
  geom_point(aes(color = species, shape= species),
             size = 3,
             alpha = 0.8) +
  theme_minimal()
```


```{r question2c, echo = FALSE, fig.width = 5.5, fig.asp = 0.618, dev = "svg"}
```


# Question 2d: species-specific shapes and colors
d. Combine b-c.    


```{r question2d, eval = FALSE}
penguins %>% 
  ggplot() + 
  aes(x = flipper_length_mm,
      y = body_mass_g) +
  geom_point(aes(color = species,
                 shape = species),
             size = 3,
             alpha = 0.8) +
  theme_minimal()
```


```{r question2d, echo = FALSE, fig.width = 5.5, fig.asp = 0.618, dev = "svg"}
```


# Question 2e: add titles

e. Change colors to darkorange, purple and cyan4. Add titles to axes, legend and figure.     


```{r question2e, eval = FALSE}
penguins %>% 
  ggplot() + 
  aes(x = flipper_length_mm,
      y = body_mass_g) +
  geom_point(aes(color = species, 
                 shape = species),
             size = 3,
             alpha = 0.8) +
  scale_color_manual(values = c("darkorange","purple","cyan4")) +
  labs(title = "Penguin body mass wrt flipper length",
       subtitle = "for Adelie, Chinstrap and Gentoo species",
       x = "Flipper length (mm)",
       y = "Body mass (g)",
       color = "Penguin species",
       shape = "Penguin species") +
  theme_minimal()
```



# Question 2e

```{r question2e, echo = FALSE, fig.width = 5.5, fig.asp = 0.818, dev = "svg"}
```


