---
title: "Reproducible science workshop"
subtitle: "Exercise 1 Tidyverse Data manipulation"
author: "[**_Gbadamassi G.O. Dossa_**](https://scholar.google.com/citations?user=w20AkZcAAAAJ&hl=en)"
date: (Created on Sept 31 2021 and last updated on `r format(Sys.time(), "%a %b %d %Y %H:%M:%S")`)
output: 
#rmdformats::material
  html_document:
    fig_caption: yes
    theme: spacelab #readable #sandstone #spacelab #flatly
    highlight: pygments #tango #kate
    toc: TRUE
    toc_depth: 3
    number_sections: TRUE
    toc_float:
      collapsed: TRUE
      smooth_scroll: TRUE
      toc: true
---

```{r setup, include=FALSE, echo=FALSE}
#options(htmltools.dir.version = FALSE)
#knitr::opts_chunk$set(comment = "", warning = FALSE, message = FALSE)
knitr::opts_chunk$set(warning = FALSE, message = FALSE)
```


# Download packages

We will need the following packages. Please run the code below to download them. 

***Please remove the "#" before running the code below***

```{r,}
#install.packages(c("tidyverse", "sf", "emo", "janitor", "palmerpenguins", "usethis", "lubridate"))
```

# Motivation

We use data from package `palmerpenguins` to familiarize ourselves with data manipulation with the tidyverse. We first load the packages we will need. 
```{r}
library(palmerpenguins)
library(tidyverse)
```


# Question 1  

a. Display the data `penguins`.  

b. Make sure you understand the columns we have in this dataset.  

c. Filter out penguins for which sex is missing.  

d. Select variables species, island, bill_length_mm and body_mass_g.  

e. Store the new dataset in a dat object.  


# Question 2  

a. How many penguins do we have in the dataset?  

b. How many species?   

c. How many islands?   

d. Count the number of penguins per species.   

e. Count the number of penguins per species and per island.    


# Question 3  

a. Calculate the overall mean body mass.  

b. Calculate the mean body mass for each species.  

c. Calculate the mean of both traits bill length and body mass measured for each species.   


